---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Valerie Chu"
date: today

format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true

execute:
  warning: false
  
from: markdown+emoji 

---

::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-1-2023-fall/final-project-1-valerieyichu.git](https://github.com/stat301-1-2023-fall/final-project-1-valerieyichu.git)

:::
## Load Packages

```{r}
library(tidyverse)
```


## Load Data

```{r}

books <- read_delim("data/books.csv")
books_tags <- read_csv("data/book_tags.csv")
ratings <- read_csv("data/ratings.csv")
tags <- read_csv("data/tags.csv")
tbr <- read_csv("data/to_read.csv")

```

## Before we dive into the data — A quick introduction to Goodreads

What is Goodreads? 

It's essentially what I think of as a combination of Twitter and Amazon for book lovers. 

These are a few of Goodread's draws. It allows users to:

- Categorize books on their bookshelves

- Write book reviews

- Find their next read

- Vote on their annual favorites across genres

- See what books their friends are reading

- Track the books they're reading, have read, and want to read

- See their personalized book recommendations. (Goodreads says "Our recommendation engine analyzes 20 billion data points to give suggestions tailored to your literary tastes.")

- Read community reviews

Officially, this is how Goodreads describes itself: "Goodreads is the world’s largest site for readers and book recommendations. Our mission is to help readers discover books they love and get more out of reading. Goodreads launched in January 2007."


## Data source

I'm interested in analyzing user habits on Goodreads. 



**First, a note about the data**

There are five datasets relevant to achieving my goal. 

- The primary dataset that I will be using is the one I have saved as "books". 

- However, the other four datasets — "books_tags", "ratings", "tags", and "tbr" all contain additional data about user habits on Goodreads. 

- The good thing is, the other four datasets can all be joined with the "books" dataset to create one large dataset. For example, `goodreads_book_id` in the "book_tags" dataset corresponds with `goodreads_book_id` in the "books" dataset. Because every book has a unique `goodreads_book_id`. This is the case across all five datasets pertaining to Goodreads user habits. 


## Why this data

Blah blah blah^[another footnote]

## Data quality & complexity check

Blah blah blah

## Potential data issues

Blah blah blah

## Misc

Blah Blah
