---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "YOUR NAME"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
---

::: {.callout-important}
## Memo format

Like all other submissions for this course, we expect the memo to be neat and well organized.

You should make this document your own. This is a template. To help with this we have added a date to the yaml above (good whenever writing memos). You should remove this block when making this document your own.

:::

::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-1-2023-fall/final-project-1-valerieyichu.git](https://github.com/stat301-1-2023-fall/final-project-1-valerieyichu.git)

:::

::: {.callout-warning}
GitHub (free account) cannot store large files. If you try to commit and push a large dataset you will have an ERROR! Any file over 100 MB (100,000 KB) needs to be added to the `.gitignore` file BEFORE committing.

**We may need to do that for files in your final project!**
:::

## Load Packages
```{r}
library(tidyverse)
```


## Load Data
```{r}

books <- read_delim("data/books.csv")
books_tags <- read_csv("data/book_tags.csv")
ratings <- read_csv("data/ratings.csv")
tags <- read_csv("data/tags.csv")
tbr <- read_csv("data/to_read.csv")

```



## Data source

Blah blah blah^[where blah blah came from, an example of footnote]

## Why this data

Blah blah blah^[another footnote]

## Data quality & complexity check

Blah blah blah

## Potential data issues

Blah blah blah

## Misc

Blah Blah
