---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Valerie Chu"
date: today

format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true

execute:
  warning: false
  
from: markdown+emoji 

---

**Data should be merged, cleaned, and variables should be explored. Univariate and bivariate analysis completed for several variables.** 
If your tibble is particularly big, put an appendix. Or if you have miscellaneous information that's not that important but that you really want to show, put it in an appendix. 

Either use ggsave or echo false. 


## Load Data
```{r}
library(tidyverse)
```



After you do all this data cleaning, save it as an RDS to preserve variable types!! That way you can read it in again with each script. 

Here's a demo of what it could look like:
```{r}
#| echo: false

tibble <- tibble(name = c("var1") , value = c(1))

tibble |> 
  knitr::kable()

```

Don't show a skim or a summary output. INSTEAD, make it look nice. 

- Ex. Calculate summary stats with summarize() to output as a nice table. Mean, sd, correlation, etc. 

- Ex. Make a scatterplot, a boxplot, etc. 


## Basic objectives
Objective 1
Students are expected to setup their own qmd file to render to an html for this project. The document should be appropriately formatted (see previous memo and other assignments). Should have a title, author, date, and appropriate headers, and sub-headers.

## Objective 2
Students are expected to demonstrate that significant progress has been made on their final project since the submission of progress memo 1. Students should have their data cleaned and the EDA should be started.

Demonstrating significant progress means students should have some univariate and bivariate analyses complete for several of their variables.  They should share a few graphics and/or tables with a description of what they have found thus far to demonstrate they progress. Students should should clearly state what they are exploring and why in these demonstrations. That is, they should share the guiding curiosity or research question that accompanies the particular graphics and/or tables they choose to share. 

## What else should be in the memo
Students should summarize their progress, where they are at, and what their next steps will be. Self assessment of progress would also be appropriate. When thinking or describing next steps students should share any guiding curiosities or research questions they plan to explore.

## Misc Notes/Comments
Final GitHub repo link should be at the top in a callout block --- similar to all other assignments
There should be no code visible or accessible in memo. 
There should be no raw R output like tibbles/data frames. Use html tables.
In rare cases where the the project is extremely heavy on advanced data collection, progress will look quite different. Projects like this will be focused on showing progress on getting the data together and collected. If your project falls in this category, then discuss this with your instructor --- very few, if any projects, fall into this rare case. 

